---
title: "CharlesCoffey17835PSET1"
author: "Charles Coffey"
date: "9/16/2020"
output:
  word_document: default
  pdf_document: default
---

QUESTION 1
```{r}
library(ISLR)
library(MASS)

#1.1 load csv files

cities_info = read.csv("cities_info.csv");
distance_to_wuhan = read.csv("distance_to_wuhan.csv");

#1.2 merged data set with all info
cities_all <- merge(cities_info, distance_to_wuhan, by = "city_name");

head(cities_all)
```
```{r}
#1.3 making the data easier to work with because it is skewed

log.wuhan.outflow <- log(cities_all$wuhan_outflow.Jan1.to.Jan24.);
log.cum.cases <- log(cities_all$cumulative_confirmed_cases.Feb19.);

#1.4 plotting outflow from wuhan vs number of cases

plot(log.wuhan.outflow, log.cum.cases, pch = 16, col = "red", 
     main = "Wuhan Outflow vs. Cummulative Number of Cases", 
     xlab = "Log of Outflow from Wuhan",
     ylab = "Log of Cummulative Cases");
```
```{r}
#1.5 plot texts instead of points
plot(log.wuhan.outflow, log.cum.cases, type="n")
text(log.wuhan.outflow, log.cum.cases, cities_all$city_name, cex = 0.45)

abline(a=-1.4193, b=0.5662)
```
```{r}
#1.6 checking for cities with potential to be seriously affected
cities.with.risk <- character();

for (row in 1:nrow(cities_all)){
  
  if ((cities_all[row, 3]*10^4 > 10^7) && (cities_all[row, 8] > 10^4)){
    
    cities.with.risk <- c(cities.with.risk, cities_all[row,1])
  }
}

print(cities.with.risk)
```

QUESTION 2
```{r}
#2.1
results_germany_2017 <- read.csv("results_germany_2017.csv", sep = ";");
print(results_germany_2017[c(3,5),])
```
Percentage of Votes of the AfD and The Greens in the First and Second Votes listed above.

```{r}
#2.2
ESS_Germany_2018 <- read.csv("ESS_Germany_2018.csv");

  # initialize zero vectors for count number of votes for each party
num_votes_1 = integer(length(results_germany_2017$party_name)) 
num_votes_2 = integer(length(results_germany_2017$party_name))

  #create holding place for keeping track of votes for each party
party_vote_estimates = data.frame(num_votes_1, num_votes_2, row.names = results_germany_2017$party_name)

  #check to see which voters voted and which parties they voted for
for (respondent in 1:nrow(ESS_Germany_2018)){
  if ((ESS_Germany_2018[respondent, "vote"] == "Yes") && (!is.na(ESS_Germany_2018[respondent, "prtvede1"]))){
    party_voted_1 = ESS_Germany_2018[respondent, "prtvede1"]
    party_vote_estimates[party_voted_1, "num_votes_1"] = party_vote_estimates[party_voted_1, "num_votes_1"] + 1
  }
  if ((ESS_Germany_2018[respondent, "vote"] == "Yes") && (!is.na(ESS_Germany_2018[respondent, "prtvede2"]))){
    party_voted_2 = ESS_Germany_2018[respondent, "prtvede2"]
    party_vote_estimates[party_voted_2, "num_votes_2"] = party_vote_estimates[party_voted_2, "num_votes_2"] + 1
  }
}

  #find percentage of voters each party received
percentage_votes_1 = party_vote_estimates$num_votes_1/sum(party_vote_estimates$num_votes_1)*100
percentage_votes_2 = party_vote_estimates$num_votes_2/sum(party_vote_estimates$num_votes_2)*100
names(percentage_votes_1) = c("Percentage Voters 1")
names(percentage_votes_2) = c("Percentage Voters 2")

  #add percentage information to storage data frame
party_vote_estimates <- cbind(party_vote_estimates, percentage_votes_1, percentage_votes_2)

  # calculated difference between expected outcome vs actual outcome 
estimated_minus_result_1 = party_vote_estimates$percentage_votes_1 - results_germany_2017$vote1_perc
estimated_minus_result_2 = party_vote_estimates$percentage_votes_2 - results_germany_2017$vote2_perc

  # add differences
party_vote_estimates <- cbind(party_vote_estimates, estimated_minus_result_1, estimated_minus_result_2)

  # create 7 column data table with all necessary information
data_table = cbind(results_germany_2017$party_name, 
                   party_vote_estimates$percentage_votes_1,
                   results_germany_2017$vote1_perc, 
                   party_vote_estimates$estimated_minus_result_1,
                   party_vote_estimates$percentage_votes_2,
                   results_germany_2017$vote2_perc,
                   party_vote_estimates$estimated_minus_result_2);

  # converting data table to data frame to be able to work with it
data_table = as.data.frame(data_table);
names(data_table) = c("Party Name", "Estimated First Vote", "Actual First Vote","Estimated1-Actual1",
                      "Estimated Second Vote", "Actual Second Vote", "Estimated2-Actual2")
  #converting all numbers to numerics instead of characters
data_table$`Estimated First Vote` = as.numeric(data_table$`Estimated First Vote`)
data_table$`Actual First Vote` = as.numeric(data_table$`Actual First Vote`)
data_table$`Estimated1-Actual1` = as.numeric(data_table$`Estimated1-Actual1`)
data_table$`Estimated Second Vote` = as.numeric(data_table$`Estimated Second Vote`)
data_table$`Actual Second Vote` = as.numeric(data_table$`Actual Second Vote`)
data_table$`Estimated2-Actual2` = as.numeric(data_table$`Estimated2-Actual2`)

  #order data table by most actual first votes
data_table[order(data_table$`Actual First Vote`),];

  #prepare data to be plotted and extract top 5 parties' information
plot_data = t(data_table[c(1:5), c(2:3)])
colnames(plot_data) <- c(data_table$`Party Name`[c(1:5)])

  #create bar plot for data
colors.names = c("cadetblue","chocolate")
par(cex.axis = 0.5)
barplot(as.matrix(plot_data),
        col = colors.names, 
        beside = TRUE,
        legend=rownames(plot_data),
        xlab = "Party",
        ylab = "Percentage",
        ylim = c(0,40),
        main = "Percentage of Voters per Party in 2017 Bundestag Election")

print(data_table[c(1:5),])
```
CONCLUSION:
The Alternative for Germany party and The Left Party gain from people lying because it is evident that they receive more voters than their survey results suggest that they would. People seem to feel ashamed that they vote for these two parties. They get more votes than they actually expect. AFD gains the most from this phenomenon. I do not think the data really suggests that German progressive voters are lazy. Parties that had less actual voters than estimated voters have a very small margin of error. This margin is not sufficient enough data to conclude that the progressive voters are lazy, except for the The Greens who had a significant margin of error. The differences seen with the parties that have higher actual voters than estimated voters suggest that these voters may perhaps be shy.

By the problem set's definition of "benefit", the Christian Democratic Union, Social Democratic Party, and the Greens benefit from people lying because they seem more popular in the polls than they truly are once votes are casted.

QUESTION 3
```{r}
#3.1
library(foreign)
LupPon_data <- read.dta("LupPon_APSR.dta");

country_names <- unique(LupPon_data$country);
years <- unique(LupPon_data$year);

redist <- na.omit(LupPon_data$redist)
ratio9050 <- na.omit(LupPon_data$ratio9050)
ratio5010 <- na.omit(LupPon_data$ratio5010)
country_3_obs_count <- 0; #count of countries with all 3 observations

for (row in 1:nrow(LupPon_data)){
  
  if (all(!is.na(LupPon_data[row, c("redist","ratio9050","ratio5010")]))){
    country_3_obs_count = country_3_obs_count + 1;
  }
}
print(paste("There are",toString(length(country_names)),"countries in the dataset.", sep = " "))
print(paste("There are",toString(length(country_names)),"years in the dataset.", sep = " "))
print("Years:")
print(years)
print(paste("There are",toString(country_3_obs_count),"country-year with all three variables in the dataset.", sep = " "))
```

```{r}
#3.2

top_ordered9050 <- LupPon_data[order(LupPon_data$ratio9050, decreasing = TRUE)[c(1:5)], c("country", "ratio9050")]

bottom_ordered9050 <- LupPon_data[order(LupPon_data$ratio9050, decreasing = FALSE)[c(1:5)], c("country", "ratio9050")]

print(top_ordered9050)
print(bottom_ordered9050)
```

```{r}

#3.3

top_ordered5010 <- LupPon_data[order(LupPon_data$ratio5010, decreasing = TRUE)[c(1:5)], c("country", "ratio5010")]

bottom_ordered5010 <- LupPon_data[order(LupPon_data$ratio5010, decreasing = FALSE)[c(1:5)], c("country", "ratio5010")]
print(top_ordered5010)
print(bottom_ordered5010)
```
Looking at this measure of inequality changes our results. This measure of inequality focuses on the divide between the middle and lower classes.
```{r}
#3.4

count1 = 0
count2 = 0

  # get rid of country-year pairs that have incomplete data
LupPon_data = LupPon_data[complete.cases(LupPon_data[c('ratio9050','ratio5010')]),]
  
skew = c()
  # calculate ratios for each country-year pair
for (row in 1:nrow(LupPon_data)){
  # variable for 9050/5010
  r9overr5 = LupPon_data[row, "ratio9050"]/LupPon_data[row, "ratio5010"]
  
  skew = append(skew, r9overr5)
  
  if(r9overr5 > 1){
    count1 = count1 + 1
  }
  else{
    count2 = count2 + 1
  }
}

  #add skew to original data frame
LupPon_data = cbind(LupPon_data, skew)

  #box plot data
par(las = 2)
par(cex.axis=0.8)
boxplot(LupPon_data$skew ~ LupPon_data[["country"]], boxwex=0.6, ylab="Ratio between 90-50 and 50-10 percentiles (skew)")

print(paste("There are", toString(count1),"country-year observations with skew>1.", sep = " "))
print(paste("There are", toString(count2),"country-year observations with skew<1.", sep= " "))
```
France has the highest skew and Canada has the lowest skew. Skews below one indicate a large amount of income disparity while those closer to one indicate less income disparity between the upper and lower classes.
